---
title: "Analysis of Dunkin' Donuts in Connecticut and New England"
author: "Andrew Ba Tran"
date: "June 24, 2015"
output: html_document
---

This analysis accompanies the TrendCT.org story: [Dunkin’ Donuts owns name rights to Hartford ball park, already owns New England](http://trendct.org/2015/06/15/dunkin-donuts-owns-rights-to-hartford-stadium-already-owns-new-england/)

Here’s the raw data we [pulled and cleaned](https://github.com/trendct/dunkin-donuts-ct/blob/master/dunkindonuts.csv) from [DunkinDonuts.com](http://www.dunkindonuts.com).

***

#### Let's start out by mapping out all the Dunkin' Donuts locations

```{r fig.width=7, fig.height=6, out.width=900,out.height=600}
library(leaflet)
dunk <- read.csv("dunkindonuts.csv", stringsAsFactors=FALSE)
leaflet(dunk) %>% addTiles('http://{s}.basemaps.cartocdn.com/dark_all/{z}/{x}/{y}.png', attribution='Map tiles by <a href="http://stamen.com">Stamen Design</a>, <a href="http://creativecommons.org/licenses/by/3.0">CC BY 3.0</a> &mdash; Map data &copy; <a href="http://www.openstreetmap.org/copyright">OpenStreetMap</a>') %>% 
  setView(-98.557087, 39.810502, zoom = 4) %>% 
  addCircles(~lng, ~lat, weight = 1, radius=3, 
                 color="#ffa500", stroke = FALSE, fillOpacity = 0.8)


```

***

#### Let's look specifically at Connecticut

***


```{r fig.width=7, fig.height=6, out.width=900,out.height=600}
library(dplyr)
library(ctnamecleaner)
library(stringr)
ct <- dunk %>% filter(state=="CT") %>% count(city)

ctnamecleaner(city, ct)

ct <- data.frame(tapply(CTDATA$n, CTDATA$real.town.name, sum))
ct$town <- rownames(ct)
colnames(ct) <- c("Dunkins", "id")

ctpopulator(id, ct)

ct <- CTPOPULATED
ct$percapita <- (ct$Dunkins/ct$pop2013)*10000
ct$percapita <- round(ct$percapita, digits=2)
ct$id <- str_to_title(ct$id)

require(gtools)
require(ggplot2)
require(rgdal)
require(scales)
require(ggmap)
require(Cairo)
require(gpclib)
require(maptools)
require(reshape)
gpclibPermit()
gpclibPermitStatus()

towntracts <- readOGR(dsn="townsmap", layer="towns")
towntracts <- fortify(towntracts, region="NAME10")

Dunk_Data <- left_join(towntracts, ct)

Dunk_Data$Dunkins <- as.numeric(Dunk_Data$Dunkins)
Dunk_Data$percapita <- as.numeric(Dunk_Data$percapita)

dd <- ggplot() +
  geom_polygon(data = Dunk_Data, aes(x=long, y=lat, group=group, 
                                         fill=percapita), color = "black", size=0.2) +
  coord_map() +
  scale_fill_distiller(type="seq", palette = "Reds", breaks=pretty_breaks(n=5)) +
  theme_nothing(legend=TRUE) +
  labs(title="Dunkin' Donuts stores per 10,000 residents", fill="")
dd

dd_all <- ggplot() +
  geom_polygon(data = Dunk_Data, aes(x=long, y=lat, group=group, 
                                         fill=Dunkins), color = "black", size=0.2) +
  coord_map() +
  scale_fill_distiller(type="seq", palette = "Reds", breaks=pretty_breaks(n=5)) +
  theme_nothing(legend=TRUE) +
  labs(title="Total Dunkin' Donuts by town", fill="")
dd_all

```

***

####Analysis for Connecticut 
```{r fig.width=7, fig.height=6, out.width=900,out.height=600}
library(DT)
ct <- ct[c("id", "pop2013", "Dunkins", "percapita")]
colnames(ct) <- c("Town", "Population", "Dunkins", "Per Capita")
datatable(ct)

```

***

### Expanding scope to New England

```{r fig.width=7, fig.height=6, out.width=900,out.height=600}
dc <- read.csv("DunkinCount.csv", stringsAsFactors=FALSE)
dc_table <- dc[c("Towns", "State", "Population", "Dunkin.Donuts","Per.1.000.residents")]
colnames(dc_table) <- c("Town", "State","Population","Dunkins", "Per Capita")
datatable(dc_table)
```

***

###Is there a correlation between a town's population and number of Dunkin' Donuts?

```{r fig.width=7, fig.height=6, out.width=900,out.height=600}

ddne <- ggplot(data=dc, aes(x=Dunkin.Donuts, y=Population)) +
  geom_point(aes(text=paste(Towns, State, sep=", ")))+ 
               geom_smooth(method=lm, formula = y ~ poly(x, 2), size=1) +
               theme_minimal() +
               ggtitle("Dunkin' Donuts versus town population in New England") +
               labs(y="Population", x="Dunkin' Donuts")
ddne

```

That outlier in the top right is Boston.

Will have to readjust by taking out the outlier and trying a polynomial model.


```{r fig.width=7, fig.height=6, out.width=900,out.height=600}


dcnotboston <- filter(dc, Towns!="Boston")

p <- ggplot(data=dcnotboston, aes(x=Dunkin.Donuts, y=Population)) +
  geom_point(aes(text=paste(Towns, State, sep=", ")))+ 
  geom_smooth(method=lm, formula = y ~ poly(x, 2), size=1) +
  theme_minimal() +
  theme(text=element_text(family="Lato", size=14)) +
  ggtitle("Dunkin' Donuts versus town population in New England") +
  labs(y="Population", x="Dunkin' Donuts")
p
```

****

### Let's apply this model to predict how many Dunkin' Donuts a town should have.
```{r fig.width=7, fig.height=6, out.width=900,out.height=600}

#Determining which factor of polynomial is most ideal
dcm_lm1 <- lm(dcnotboston$Dunkin.Donuts ~ poly(dcnotboston$Population, 1))
summary(dcm_lm1)

dcm_lm2 <- lm(dcnotboston$Dunkin.Donuts ~ poly(dcnotboston$Population, 2))
summary(dcm_lm2)

dcm_lm3 <- lm(dcnotboston$Dunkin.Donuts ~ poly(dcnotboston$Population, 3))
summary(dcm_lm3)

#OK, looks like the ideal is 2

new <- data.frame(Population = dcnotboston$Population)

ddpredict <- predict(dcm_lm2, newdata = new, interval="confidence")

predicted <- cbind(dcnotboston, ddpredict)
predicted <- predicted %>% filter(State=="CT")
predicted$rounded <- round(predicted$upr, digits=0)
predicted$prediction <-predicted$rounded - predicted$Dunkin.Donuts
predicted <- predicted[order(predicted$prediction),]

total <- predicted[c("Towns", "Dunkin.Donuts", "Population", "fit", "lwr", "upr", "rounded", "prediction")]

total$fit <- round(total$fit, digits=2)
total$lwr <- round(total$lwr, digits=2)
total$upr <- round(total$upr, digits=2)

datatable(total)

```



